---
title: "Purple Urchins Contorl vs Annual"
author: "Russell Fager"
date: "10/23/2021"
output: html_document
---

## Overview

This code wrangles then visualizes data for purple urchins observed at the Naples site in 2010.

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

```{r setup, include=TRUE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)

```

## Read in the data

read in the urchins.csv data

```{r}
urchins_data <- read_csv(here("data", "urchins.csv"))
```

## Clean names to lower snake case

```{r}
urchins_data_clean <- urchins_data %>%  clean_names()
```

## Data Wrangling

The below code wrangles the urchins data to only include Naples Reef observations for purple urchins in 2010

```{r}
purple_urchins <- urchins_data_clean %>% 
  filter(common_name == "Purple Urchin", site == "NAPL") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  filter(year == "2010")

```

## Data Visualization

The code below creates a histogram comparing the size of purple urchins in each treatment group (control vs. annual)

```{r}
ggplot(data = purple_urchins, aes(x = size, color = treatment)) +
  geom_histogram() +
  xlim(0, 10)+
  facet_wrap( ~ treatment)
 
```

The code below creates a qqplot of purple urchin size from the Naples Reef observations in 2010

```{r}
qqnorm(purple_urchins$size)
qqline(purple_urchins$size)
```

The data points are close to the line indicating that the data has a close to normal distribution. This is also mirrored in the histogram which shows that the spread of purple urchin size in both treatments is very similar

## Descriptive Statistics Table

Below is a basic summary table of descriptive statistics for both groups

```{r}
purpl_urchin_stats <- purple_urchins %>% 
  group_by(treatment) %>% 
  summarize(
    n = n(),
    mean = mean(size),
    sd = sd(size),
    median = median(size)) %>% 

kable(caption = "Summary Statistics of Purple Urchin Size by Treatment", digits = 2, col.names = c("species",
                       "Sample Size",
                       "Mean Size (cm)",
                       "Standard Deviation (cm)",
                       "Median Size (cm)")) %>% 
  kable_styling(full_width = FALSE)
purpl_urchin_stats

```

## End Task
